---
title: "R Notebook"
output: html_notebook
---

# 1. Introduction

Analyses file for the `CodeClan - dirty data project - task 5: Right Wing Authoritarianism
Thijmen Breeschoten 
CodeClan Professional Data Analysis student cohort DE21

# Analysis


```{r}
library(tidyverse)

rightwing <- read_csv("../clean_data/rightwing_cleaned.csv") 
```

## Question 1 - What’s the average RWA score for each gender?

```{r}
rightwing %>% 
  select(gender, average_RWS) %>% 
  group_by(gender) %>% 
  summarise(average_RWA_per_gender = mean(average_RWS))
```

## Question 2 - What’s the average RWA score for left handed people vs. right handed people

```{r}
rightwing %>% 
  filter(hand == "Right" | hand == "Left") %>% 
  group_by(hand) %>% 
  summarise(average_RWA_per_handedness = mean(average_RWS))
```

## Question 4 - What’s the average time to take the test for each education level?

```{r}
time_per_edulevel <- rightwing %>% 
  select(testelapse, education) %>% 
  group_by(education) %>% 
  summarise(average_test_time_per_edu_level = mean(testelapse)) 

time_per_edulevel
```

## Question 6 - What’s the average RWA score for people aged Under 18, 18 to 25, 26 to 40, 41 to 60, Over 60?

```{r}
rightwing %>% 
  select(age_groups, average_RWS) %>% 
  group_by(age_groups) %>% 
  summarise(average_RWA_per_age_group = mean(average_RWS))
```

