---
title: "R Notebook"
output: html_notebook
---

# CodeClan dirty data project: sea bird observation
# this script contains code to analyse the cleaned data set 'seabirds_cleaned.csv'
# The cleaned dataset is cleaned using the script 'cleaning_seabirds.R'

##Analyses question 1
Which bird had the most individual sightings?
  count the total observations for all birds, only for count=1
  
```{r}
library(tidyverse)
seabirds <- read_csv("../clean_data/seabirds_cleaned.csv")

head(seabirds)
names(seabirds)
```
```{r}
seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  filter(count == 1) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(count_birds_individual_sights = sum(count)) %>% 
  arrange(desc(count_birds_individual_sights))
```
### Answer:
Species with the highest individual counts is the species complex:
Diomedea antipodensis / exulans, Wandering albatross sensu lato, DIOANTEXU
-> 6508 exx.

If we want to have the highest scoring true species:
Pterodroma macroptera gouldi / Grey-faced petrel / PTEMACGOU
-> 1023 exx.

##Analyses question 2
Which bird had the highest total count?
```{r}
seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  
```

