---
title: "R Notebook"
output: html_notebook
---

# CodeClan dirty data project: sea bird observation
# this script contains code to analyse the cleaned data set 'seabirds_cleaned.csv'
# The cleaned dataset is cleaned using the script 'cleaning_seabirds.R'

##Analyses question 1
Which bird had the most individual sightings?
  count the total observations for all birds, only for count=1
  
```{r}
library(tidyverse)
seabirds <- read_csv("../clean_data/seabirds_cleaned.csv")

head(seabirds)
names(seabirds)
```
```{r}
seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  filter(count == 1) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(count_birds_individual_sights = sum(count)) %>% 
  arrange(desc(count_birds_individual_sights))
```
### Answer:
Species with the highest individual counts is the species complex:
Diomedea antipodensis / exulans, Wandering albatross sensu lato, DIOANTEXU
-> 6508 exx.

If we want to have the highest scoring true species:
Pterodroma macroptera gouldi / Grey-faced petrel / PTEMACGOU
-> 1023 exx.

##Analyses question 2
Which bird had the highest total count?
```{r}
seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  head(1)
```
### Answer:
Puffinus tenuirostris - Short-tailed shearwater - PUFTEN - total count: 982553 sights!

##Analyses question 3
Which bird had the highest total count above a latitude of -30?

-Note: Here I assume above a latitude of -30, that we look at counts done North of 
      the -30 latitude (towards the equator = 0)

```{r}
seabirds %>% 
  filter(lat > -30) %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(total_count = sum(count)) %>% 
  arrange(desc(total_count)) %>% 
  head(1)
```
### Answer:
Puffinus pacificus - Wedge-tailed shearwater - PUFPAC - total count: 855

##Analyses question 4
How many different types of birds were only ever seen in groups of 1?

-> For this question I only want to look at observations on species level only.
  I will filter out all 'species complexes', 'sensu lato's' and identifications 
  on genus level.

```{r}
#-> For this question I only want to look at observations on species level only.
#  I will filter out all 'species complexes', 'sensu lato's' and identifications 
#  on genus level. (hence: triple filter)

seabirds %>% 
  filter(count == 1) %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>%
  filter(!str_detect(species_common, "sensu lato")) %>% 
  filter(str_detect(species_scientific, "[A-Z]{1}[a-z]* [a-z]*$")) %>% 
  filter(!str_detect(species_common, "unidentified")) %>% 
  filter(!str_detect(species_scientific, "\\/")) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(count_birds_individual_sights = sum(count)) %>% 
  arrange(count_birds_individual_sights) %>% 
  filter(count_birds_individual_sights == 1) %>% 
  nrow()

```
### Answer:
a total of 8 unique bird species (truly identified to species level) were ever only 
observed in singletons.

##Analyses question 5
How many penguins were seen? (Hint: there are many types of penguin)

```{r}
# there are 6 penguin genera: 
# Aptenodytes, Pygoscelis, Eudyptula, Spheniscus, Megadyptes, Eudyptes

seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  filter(str_detect(species_scientific, "Aptenodytes|Pygoscelis|Eudyptula|Spheniscus|Megadyptes|Eudyptes")) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(count_penguins_per_species = sum(count)) %>% 
  arrange(count_penguins_per_species)  

# In total 6 species (of which one species complex)

seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  filter(str_detect(species_scientific, "Aptenodytes|Pygoscelis|Eudyptula|Spheniscus|Megadyptes|Eudyptes")) %>% 
  summarise(count_total_penguin_individuals = sum(count)) 

# 154 individual penguins! 
  
  

```
### Answer:
In total 6 species (of which one species complex)
In total 154 penguin individuals
