---
title: "README"
author: "Thijmen Breeschoten"
date: "2023-06-12"
output: html_document
---

# README 

README file for the `CodeClan - dirty data project - task 3: seabirds observations
Thijmen Breeschoten 
CodeClan Professional Data Analysis student cohort DE21
This file does provide information as normally presented in a README file, but 
further sums up information regarding the cleaning/analyses steps taken for review.

### Overview
Goal of the project task is to clean and analyse a raw dataset in order to answer 
several questions regarding the data.
The raw data contains seabird observations/sightings as done from ships in seas 
around New Zealand. The raw data (.xls file) contains two worksheets with 1) ship data 
and 2) bird data (counts+names). 

The underlying code provides:
1. data cleaning steps
2. code to combine ship data (with dates, location, etc) with bird information
  (species, counts).
3. data analyses

### Language, software and packages used

* The code is written in R (version 4.3.0)
  R Core Team (2023). _R: A Language and Environment for Statistical Computing_. 
  R Foundation for Statistical Computing, Vienna, Austria.
   <https://www.R-project.org/>

* R is run in Rstudio 2023.03.0+386 

* The following libraries are required:
  cleaning_seabirds.R
  - `tidyverse`(ver 2.0.0)
  - `readxl` (ver 1.4.2)
  analysing_seabirds.Rmd
  - `tidyverse`(ver 2.0.0)
  
### File and folder overview
The project folder `sea_bird_observation` contains the .Rproj file and this README file.
The following files are included and distributed in each folder:

clean_data
  - `seabirds_cleaned.csv` cleaned dataset using script 'cleaning_seabirds.R'
data_cleaning_scripts
  - `cleaning_seabirds.R` Rscript to read and clean the raw data file.
documentation_and_analysis
  - `analysing_seabirds.Rmd` (and knitted html file) Rnotebook script to analyse the
    cleaned dataset in order to answer the analyses questions.
raw_data
  - `seabirds.xls` raw dataset in xls format

### cleaning overview

* Cleaning the raw dataset using the script `cleaning_seabirds.R`
  - Step 1, read in data from both worksheets and clean names
  - Step 2, non-informative columns are dropped
  - Step 3, datasets from both worksheet are joined using a left_join 
    (since you want to keep all bird sightings info rows)
```{r}
# joining data
seabirds_observed <- left_join(seabirds_renamed, ships_renamed, "record_id")
```
  - Step 4, dropping all additional information about the age in the species name columns
```{r}
seabirds_correct_name <- seabirds_observed %>% 
  mutate(species_common = 
           str_remove(species_common, " (AD|JUV|IMM|SUBAD|PL[0-9])$")) %>% 
  mutate(species_abbreviation = 
           str_remove(species_abbreviation, " (AD|JUV|IMM|SUBAD|PL[0-9])$")) %>%
  mutate(species_scientific = 
           str_remove(species_scientific, " (AD|JUV|IMM|SUBAD|PL[0-9])$"))
```
  - Step 5, check missing data and drop NAs for all rows where no scientific name is given.
  We at least need a scientific name to know which bird is seen!
```{r}
#check NAs for important columns
seabirds_correct_name %>% 
  select(species_common, species_scientific, species_abbreviation, count, lat, long) %>% 
  summarise(across(.fns = ~sum(is.na(.x))))

# most NAs found in scientific name (if scientific name is absent, than 
#other names are not valid)
seabirds_correct_name %>% 
  select(species_common, species_scientific, species_abbreviation, count, lat, long) %>% 
  filter(is.na(species_scientific))
# therefore, lets drop all rows with NA scientific name
seabirds_names_complete <- seabirds_correct_name %>% 
  drop_na(species_scientific)
```
  -> Important assumption/ decision:
  I have not removed double names/ and or Sensu Lato descriptions from the
  observations list. In biology it is common not to be able to identify down to species
  but have a 'species complex' as a valid observation instead. If I decide to remove 
  one of the 'potential' species names I do increase the certainty about the observation,
  while it is clear that the observer could not identify the seabird down to a 
  single species. This is important information to keep in your data file!
  - Similar for species for which only a genus name is given. I made the assumption 
  that the observer could only identify to genus level, not the species.
  
  - Step 6, write cleaned dataset to new csv file
```{r}
# Write the cleaned dataset to new csv file:
  write.csv(seabirds_names_complete, file = "clean_data/seabirds_cleaned.csv")
```
  

### Assumptions / decisions
1. -> Important assumption/ decision:
  I have not removed double names/ and or Sensu Lato descriptions from the
  observations list. In biology it is common not to be able to identify down to species
  but have a 'species complex' as a valid observation instead. If I decide to remove 
  one of the 'potential' species names I do increase the certainty about the observation,
  while it is clear that the observer could not identify the seabird down to a 
  single species. This is important information to keep in your data file!
  
- Similar for species for which only a genus name is given. I made the assumption 
  that the observer could only identify to genus level, not the species. This is 
  still a valid observation to leave in the dataset. But if a species name is asked
  as question I will not take these into account.
-> I will explain the answer of each analyses question based on the assumption above.

### analyses overview
* Question 1 - Which bird had the most individual sightings?
  
```{r}
seabirds %>% 
  select(species_scientific, species_common, species_abbreviation, count) %>% 
  filter(count == 1) %>% 
  group_by(species_scientific, species_common, species_abbreviation) %>% 
  summarise(count_birds_individual_sights = sum(count)) %>% 
  arrange(desc(count_birds_individual_sights))
```
  Answer:
  Species with the highest individual counts is the species complex:
  Diomedea antipodensis / exulans, Wandering albatross sensu lato, DIOANTEXU
  -> 6508 exx.

  If we want to have the highest scoring true species:
  Pterodroma macroptera gouldi / Grey-faced petrel / PTEMACGOU
  -> 1023 exx.
  

  