---
title: "R Notebook"
output: html_notebook
---

# 1. Introduction

Analyses file for the `CodeClan - dirty data project - task 4: Halloween Candy
Thijmen Breeschoten 
CodeClan Professional Data Analysis student cohort DE21
This file does provide information regarding the cleaning/analyses steps taken for 
the 'Dirty Data' project Halloween Candy.

# 2. Overview

Goal of the project task is to clean and analyse a raw dataset in order to answer 
several questions regarding the data.
The raw data contains results from a questionnaire regarding halloween/trick or treat candy.
The questionnaires are taken in three consecutive years (2015, 2016, 2017), and
the results can be found in three separate .xlsx files:
- `boing-boing-candy-2015.xlsx`
- `boing-boing-candy-2016.xlsx`
- `boing-boing-candy-2017.xlsx`

Contents of this file:
1. Introduction
2. Overview
3. Language, software and packages used
4. File and folder overview
5. cleaning overview
6. Assumptions / decisions
7. analyses overview & answers

# 3. Language, software and packages used

* The code is written in R (version 4.3.0)
  R Core Team (2023). _R: A Language and Environment for Statistical Computing_. 
  R Foundation for Statistical Computing, Vienna, Austria.
   <https://www.R-project.org/>

* R is run in Rstudio 2023.03.0+386 

* The following libraries are required (specified per script):
  cleaning_candy.R
  - `tidyverse`(ver 2.0.0)
  - `readxl` (ver 1.4.2)
  analysing_candy.Rmd
  - `tidyverse`(ver 2.0.0)
  
# 4. File and folder overview
The Rproject folder `halloween_candy` contains the .Rproj file and the following 
folders.
The following files are included and distributed in each folder:

clean_data
  - `candy_cleaned.csv` cleaned dataset using script 'cleaning_candy.R'
data_cleaning_scripts
  - `cleaning_candy.R` Rscript to read and clean the raw data file.
documentation_and_analysis
  - `analysing_candy.Rmd` (and knitted html file) Rnotebook script presenting an 
  overview and background information regarding the project and analyses. It further   presents the script/code to analyse the cleaned dataset in order to answer the       analyses questions.
raw_data
  - ``boing-boing-candy-2015.xlsx`, `boing-boing-candy-2016.xlsx`,                       `boing-boing-candy-2017.xlsx` raw datasets in xlsx format.
  
# 5. cleaning overview

# 6. Important assumptions/decisions

# 7. Analyses
```{r}
# open libraries and cleaned file
library(tidyverse)
candy <- read_csv("../clean_data/candy_cleaned.csv")
#names(candy)
head(candy)
#glimpse(candy)
```
## Question 1 - What is the total number of candy ratings given across the three years. 
(Number of  candy ratings, not the number of raters. Don’t count missing values)

```{r}
candy %>% 
  filter(rating != is.na(rating)) %>% 
  nrow()
```
### Answer:
The total number of ratings (either meh, depspair or joy) for candy products is 524165

## Question 2 - What was the average age of people who are going out trick or treating?

```{r}
candy %>% 
  filter(participate == "Yes") %>% 
  summarise(average_age_participating = mean(age, na.rm = TRUE))
```
### Answer:
The average age of people participating is 35.22 years

## Question 3 - What was the average age of people who are not going trick or treating

```{r}
candy %>% 
  filter(participate == "No") %>% 
  summarise(average_age_NOT_participating = mean(age, na.rm = TRUE))
```
### Answer:
The average age of people NOT participating is 39.26 years

## Question 4 - For each of joy, despair and meh, which candy bar received the most of these ratings
-1. first important step is to indicate which candy types are actually bars,
    we need to do some recoding 
```{r}
#first important step is to indicate which candy types are actually bars,
#we need to do some recoding 
#(add additional column indicating if we are dealing with a candy bar) :

candy_bars <- candy %>% 
  mutate(is_bar = if_else(
    str_detect(
      candy_type, "bar") |
      candy_type %in% 
      c("hersheys_dark_chocolate", "coffee_crisp", "twix", "three_musketeers",
        "milky_way", "nestle_crunch", "butterfinger", "caramellos", "snickers",    
        "hershey_s_milk_chocolate", "kit_kat", "mars", "reese_s_peanut_butter_cups"), 
    TRUE, FALSE)
    ) %>% 
  filter(is_bar == TRUE) 
```
-2. now using the new object, we can check which bar received most ratings for
    Joy, Despair, and Meh

```{r}
# First check for Joy:
candy_bars %>% 
  select(candy_type, rating) %>% 
  filter(rating == "JOY") %>% 
  group_by(candy_type) %>% 
  summarise(total_joys_per_candy_type = n()) %>% 
  arrange(desc(total_joys_per_candy_type)) %>% 
  head(1)
```
```{r}
# Secondly, check for despair:
candy_bars %>% 
  select(candy_type, rating) %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candy_type) %>% 
  summarise(total_joys_per_candy_type = n()) %>% 
  arrange(desc(total_joys_per_candy_type)) %>% 
  head(1)
```

```{r}
# Thirdly, check for meh:
candy_bars %>% 
  select(candy_type, rating) %>% 
  filter(rating == "MEH") %>% 
  group_by(candy_type) %>% 
  summarise(total_joys_per_candy_type = n()) %>% 
  arrange(desc(total_joys_per_candy_type)) %>% 
  head(1)
```
### Answer:
For each of joy, despair and meh, highest rating candy is:
  Joy     -> reese_s_peanut_butter_cups     - 7369 votes
  Meh     -> 100_grand_bar                  - 1307 votes
  despair -> reggie_jackson_bar             - 4506 votes

## Question 5 - How many people rated Starburst as despair
```{r}
candy %>% 
  filter(candy_type == "starburst") %>% 
  filter(rating == "DESPAIR") %>% 
  summarise(count_total_despair_for_starburst = n())
```
### Answer:
In total 1990 people voted "despair" for starburst

*For the next three questions: count despair as -1, joy as +1, and meh as 0*
Lets create an additional column using the given coding system
```{r}
# All remaining questions relate to candy_bars only so do this for the dataset 
# `candy_bars`

candy_bars_vote_count <- candy_bars %>% 
  mutate(vote = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    rating == "MEH" ~ 0
  )) %>% 
  drop_na(vote)
```

## Question 6 - What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
candy_bars_vote_count %>% 
  select(gender) %>% 
  distinct(gender) %>% 
  head() 
# THere are 4 gender types: Male, Female, Other, I'd rather not say

# Lets check Male
candy_bars_vote_count %>% 
  filter(gender == "Male") %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
```{r}
# Lets check Female
candy_bars_vote_count %>% 
  filter(gender == "Female") %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
```{r}
# Lets check Other
candy_bars_vote_count %>% 
  filter(gender == "Other") %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
```{r}
# Lets check I'd rather not say
candy_bars_vote_count %>% 
  filter(gender == "I'd rather not say") %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
### Answer:
The most popular candy bar by this rating system per gender is:
- Male:                 reese_s_peanut_butter_cups	- 1443
- Female:               reese_s_peanut_butter_cups  - 768 votes
- Other:                twix        - 32 votes
- I'd rather not say:   kit_kat     - 62 votes

## Question 7 - What was the most popular candy bar in each year?

```{r}
#check 2015 first
candy_bars_vote_count %>% 
  filter(year == 2015) %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
```{r}
candy_bars_vote_count %>% 
  filter(year == 2016) %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
```{r}
candy_bars_vote_count %>% 
  filter(year == 2017) %>% 
  group_by(candy_type) %>% 
  summarise(total_vote_count = sum(vote)) %>% 
  arrange(desc(total_vote_count)) %>% 
  head(1)
```
### Answer:
The most popular candy bar by this rating system per year is:
- 2015: reese_s_peanut_butter_cups    - 4375
- 2016: kit_kat                       - 920
- 2017: reese_s_peanut_butter_cups    - 1403


```{r}
candy %>% 
  distinct(candy_type) %>% 
  print(200)
```

